NumLin: Linear Types for Linear Algebra

### 2018-11-15 write-up/paper/discussion.tex
Start Performance Metrics
Start Related Work

\subsection{Finding Bugs in SymPy's Output}

@@

The main point we want to make is that using linear types for BLAS
is an ``obvious'' idea, but is surprisingly under-explored.

\begin{itemize}
\item Rust
\item ATS
\item Single-assignment C 
\item Linear Haskell
\item Bernardy and Sveningsson
\item L3
\item Boyland fractional permissions
\item Disadvantage of graph-based approaches to matrix expression generation
\item multi-stage: once dimensions known, they are fixed?
\end{itemize}

>>>>

\subsection{Finding Bugs in SymPy's Output}\label{subsec:finding_bugs}

@@

% The main point we want to make is that using linear types for BLAS
% is an ``obvious'' idea, but is surprisingly under-explored.

Using linear types for BLAS routines is a particularly good domain fit (given
the implicit restrictions on aliasing arguments), but is under-explored in the
academic literature. Linear algebra libraries written in Rust take advantage of
the distinction that Rust's type system offers between mutable views/references
to arrays, offering a \emph{seemingly} different approach to the
fractional-permissions based one we took in this paper.  However, recent
research\cite{weiss} suggests that Rust's borrow-checker could be explained in
simpler terms using \emph{fractional-permissions}; in this light, our paper
neatly bridges the gap by demonstrating the theoretical and practical
simplicity of this way of thinking.

ATS\cite{cui2005ats} is a language which combines dependent and linear types;
although it provides BLAS bindings, the main focus in the API provided there is
for statically enforcing dimensions related constraints of the arguments,
rather than aliasing restrictions as shown here.

\begin{itemize}
\item Single-assignment C: for hardware, doesn't allow higher-order fun.
\item Linear Haskell: linearity on the arrows.
\item Bernardy and Sveningsson: very theoretical, no implementation.
\item L3: can use existential types to statically track pointer identities to
    prevent that source of runtime errors.
\item Boyland fractional permissions
\item Disadvantage of graph-based approaches to matrix expression generation:
    buggy, like SymPy.
\item multi-stage: once dimensions known, they are fixed, so we can take
    dimension checking one-step further and implement a matrix chaining alg.
\end{itemize}

% Dimensions: easy enough to integrate, even w/o ATS/dep types
% generative phantom types
% \cite{abe2015simple}
